Modal Verification Patterns for Systems

### 2025-06-17 paper.tex
+

Write amplification traverses the overall tree depth-first, transforming the tree from the bottom up, one full subtree at a time. Each time a node is split, it results in disconnecting some nodes from the next snapshot, sharing other nodes, and linking some newly-allocated nodes into the new snapshot, as in Figure \ref{fig:cow_index}. Nodes are split (or deemed not to require splitting) one at a time, and considered only after all child nodes have been amplified. Concepually this means the amplification code's proof must track the fact that at any point in time, there are regions of the tree where each node has already been amplified (e.g., $\phi_{\textsf{amplified}}$ definitely holds for that node), and regions where nodes may or may not require amplification, and this will not be known until those nodes are visited (i.e., some basic validity $\phi_{\textsf{valid}}$ is known to hold, but not the stronger $\phi_{\textsf{amplified}}$).

%\setlength\intextsep{0pt}
\begin{wrapfigure}{r}{0.25\textwidth}
  \centering  \includegraphics[width=0.25\textwidth]{entail.png}
  \caption{Understanding the level of $\varphi$ satisfaction in the \textsc{CoW} tree of modal nodes}
   \label{fig:nodeentail} 
\end{wrapfigure}
We use a modality similar to $\text{\faSitemap}@[\ell]\phi$ to indicate that $\phi$ is true of the tree node rooted at $\ell$, as well as at \emph{all reachable children} of that node.
This means that establishing a particular $\phi$ throughout a tree must be done in a bottom-up manner (see Figure \ref{fig:nodeentail}).
Capturing this modally, rather than with more typical ad-hoc hand-crafted invariants, offers advantages in a number of proof steps, in addition to concisely capturing intuitive reasoning about subtrees with all nodes satisfying a particular property. One basic example is that while the actual disk sync code requires the full filesystem tree to be in an amplified state (hence write amplification), most of the filesystem works with more general trees. Since $\phi_{\textsf{amplified}}\vdash\phi_{\textsf{valid}}$, using our \faSitemap modality we can conclude $\text{\faSitemap}@[\ell]\phi_{\textsf{amplified}}\vdash\text{\faSitemap}@[\ell]\phi_{\textsf{valid}}$ by an analogue of Axiom K.
There are also other places in the code where a subtree satisfying a stronger assertion at all nodes must be linked into a larger tree satisfying only a weaker assertion overall: predicate representing the valid tree as the tree of nodes which are backed by resources allocated pages, i.e. $\phi_{\textsf{valid}}$, is weaker than the predicate representing the tree of nodes being currently amplified in the current snapshot. Again, the same reasoning principle makes this straightforward. As alluded to earlier, our $\phi_{\textsf{amplified}}$ and company are in fact \emph{predicates} parameterized by a choice of node, though by lifing \textsf{Iris}'s BI structure they can use all of the standard separation logic connectives. The rest of this subsection develops this modality in full detail.
%\colinsaysinl{I cut some text here about the three-way split (commented out) that I wasn't sure how to use, but maybe I deleted something important?}
%To put it concretely, as shown in Figure \ref{fig:cow_index_tree.JPG}, \textsc{CoW} semantics determine what we can claim on the structure of the filesystem indexing tree:
%\begin{itemize}
%\item The bottom-up direction splits the tree into the two half at the cursor node: 1. the bottom part accessible from the cursor node and per-node-amplification is reflected in case needed 2. the upper part on which amplifcation has not reflected yet. 
%\item Horizontally, amplifcation path splits the bottom part of the tree into three different contexts: untouched tree nodes, nodes on the amplification path, and the ones generated as part of node splitting. 
%\end{itemize}
%Based on these two observations, in fact, bottom-up amplification determines the choice of modal context by giving the boundries of \textit{any claimable knowledge} (e.g. $\varphi_{\textsf{amplified}}$, $\varphi_{\textsf{untouched}}$): a subtree bounded by the cursor node.
\subsection{Nominals for Systems Resource Contexts}
Finally, another design point is the question of whether or not resource element assertions must explicitly track
their corresponding context, or if they implicitly pick up their context from where they are used.

>>>>

Write amplification traverses the overall tree depth-first, transforming the tree from the bottom up, one full subtree at a time. Concepually this means the amplification code's proof must track the fact that at any point in time, there are regions of the tree where each node has already been amplified (e.g., $\phi_{\textsf{amplified}}$ definitely holds for that node), and regions where nodes may or may not require amplification, and this will not be known until those nodes are visited (i.e., some basic validity $\phi_{\textsf{valid}}$ is known to hold, but not the stronger $\phi_{\textsf{amplified}}$).

%\setlength\intextsep{0pt}
\begin{wrapfigure}{r}{0.25\textwidth}
  \centering  \includegraphics[width=0.25\textwidth]{entail.png}
  \caption{Understanding the level of $\varphi$ satisfaction in the \textsc{CoW} tree of modal nodes}
   \label{fig:nodeentail} 
\end{wrapfigure}
We use a modality similar to $\text{\faSitemap}@[\ell]\phi$ to indicate that $\phi$ is true of the tree node rooted at $\ell$, as well as at \emph{all reachable children} of that node.
This means that establishing a particular $\phi$ throughout a tree must be done in a bottom-up manner (see Figure \ref{fig:nodeentail}).
Capturing this modally, rather than with more typical ad-hoc hand-crafted invariants, offers advantages in a number of proof steps, in addition to concisely capturing intuitive reasoning about subtrees with all nodes satisfying a particular property. One basic example is that while the actual disk sync code requires the full filesystem tree to be in an amplified state (hence write amplification), most of the filesystem works with more general trees. Since $\phi_{\textsf{amplified}}\vdash\phi_{\textsf{valid}}$, using our \faSitemap modality we can conclude $\text{\faSitemap}@[\ell]\phi_{\textsf{amplified}}\vdash\text{\faSitemap}@[\ell]\phi_{\textsf{valid}}$ by an analogue of Axiom K.
There are also other places in the code where a subtree satisfying a stronger assertion at all nodes must be linked into a larger tree satisfying only a weaker assertion overall: predicate representing the valid tree as the tree of nodes which are backed by resources allocated pages, i.e. $\phi_{\textsf{valid}}$, is weaker than the predicate representing the tree of nodes being currently amplified in the current snapshot. Again, the same reasoning principle makes this straightforward. As alluded to earlier, our $\phi_{\textsf{amplified}}$ and company are in fact \emph{predicates} parameterized by a choice of node, though by lifing \textsf{Iris}'s BI structure they can use all of the standard separation logic connectives. 
%\colinsaysinl{I cut some text here about the three-way split (commented out) that I wasn't sure how to use, but maybe I deleted something important?}
%To put it concretely, as shown in Figure \ref{fig:cow_index_tree.JPG}, \textsc{CoW} semantics determine what we can claim on the structure of the filesystem indexing tree:
%\begin{itemize}
%\item The bottom-up direction splits the tree into the two half at the cursor node: 1. the bottom part accessible from the cursor node and per-node-amplification is reflected in case needed 2. the upper part on which amplifcation has not reflected yet. 
%\item Horizontally, amplifcation path splits the bottom part of the tree into three different contexts: untouched tree nodes, nodes on the amplification path, and the ones generated as part of node splitting. 
%\end{itemize}
%Based on these two observations, in fact, bottom-up amplification determines the choice of modal context by giving the boundries of \textit{any claimable knowledge} (e.g. $\varphi_{\textsf{amplified}}$, $\varphi_{\textsf{untouched}}$): a subtree bounded by the cursor node.
\subsection{Nominals for Systems Resource Contexts}
Finally, another design point is the question of whether or not resource element assertions must explicitly track
their corresponding context, or if they implicitly pick up their context from where they are used.
